SSwAB-IT Phase 1 Analysis

### 2025-04-01 Jordana Hoppe/SSwAB-IT/phase_1_analysis.qmd
SSwAB-IT FEV1 by age

```{r setup}
#| include: false
library(redcapAPI)
library(tidyverse)
library(gt)
library(gtsummary)
library(caret)
library(eulerr)
library(ggpubr)
```

@@

  db_clinic_obtained_values %>% select(record_id, cos_collection_date),

@@

@@

@@

```{r}
db_demographics %>%
  select(group, age_at_cos, age_group, sex, Race, Ethnicity) %>%
  tbl_summary(
    by = group, missing_text = "Missing",
    statistic = all_continuous() ~ c("{median} ({min}, {max})"),
    label = list(
      "sex" = "Sex", "age_at_cos" = "Age at COS", "age_group" = "Age Group"
    )
  ) %>%
  add_overall() %>%
  add_p(
    test = list(c("Race", "Ethnicity") ~ "fisher.test"),
    test.args = age_group ~ list(workspace = 2e9)
  ) %>%
  separate_p_footnotes() %>%
  modify_footnote(everything() ~ NA) %>%
  as_gt()
```

@@

```{r}
# All pathogens
d_overall <- df[, c("overall_cos", "overall_sos")]
colnames(d_overall) <- c("COS", "SOS")
d_overall <- d_overall[complete.cases(d_overall), ]
a1 <- plot(euler(d_overall == "Positive"), quantities = list(cex = 3))
a2 <-
  ggplot(as.data.frame(table(d_overall$COS, d_overall$SOS)), aes(Var1, Var2)) +
  geom_tile(color = "white", fill = "#1E88E5") +
  geom_text(
    aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
    color = "white"
  ) +
  xlab("COS") +
  ylab("SOS") +
  theme_bw() +
  theme(text = element_text(size = 15))
# MRSA
d_MSSA <- df[, c("staphylococcus_aureus_cos", "staphylococcus_aureus_sos")]
colnames(d_MSSA) <- c("COS", "SOS")
d_MSSA <- d_MSSA[complete.cases(d_MSSA), ]
b1 <- plot(euler(d_MSSA == "Positive"), quantities = list(cex = 3))
b2 <-
  ggplot(as.data.frame(table(d_MSSA$COS, d_MSSA$SOS)), aes(Var1, Var2)) +
  geom_tile(color = "white", fill = "#1E88E5") +
  geom_text(aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
    color = "white"
  ) +
  xlab("COS") +
  ylab("SOS") +
  theme_bw() +
  theme(text = element_text(size = 15))
p1 <- ggarrange(a1, a2, b1, b2, nrow = 2, ncol = 2) +
  bgcolor("White")
ggsave("/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/fig.png",
  plot = p1, width = 9, height = 6, units = "in"
)
a1 <- plot(euler(d_overall == "Positive"),
  quantities = list(cex = 3),
  main = "Any CF Pathogen"
)
b1 <- plot(euler(d_MSSA == "Positive"),
  quantities = list(cex = 3),
  main = "MSSA"
)
p2 <- ggarrange(a1, b1, nrow = 1, ncol = 2) +
  bgcolor("White")
ggsave("/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/euler_bw.png",
  plot = p2, width = 9, height = 6, units = "in"
)
a1 <- plot(venn(d_overall == "Positive"),
  quantities = list(cex = 3),
  main = "Any CF Pathogen"
)
b1 <- plot(venn(d_MSSA == "Positive"),
  quantities = list(cex = 3),
  main = "MSSA"
)
p3 <- ggarrange(a1, b1, nrow = 1, ncol = 2) +
  bgcolor("White")
ggsave("/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT/Dissemination/Figures/venn_bw.png",
  plot = p3, width = 9, height = 6, units = "in"
)
```

>>>>

```{r setup}
#| include: false
library(redcapAPI)
library(tidyverse)
library(gt)
library(gtsummary)
library(caret)
library(eulerr)
library(ggpubr)
home_dir <- switch(Sys.info()["sysname"],
  "Darwin" = "/Users/timvigers/Library/CloudStorage/OneDrive-TheUniversityofColoradoDenver/Vigers/CF/Jordana Hoppe/SSwAB-IT",
  "Linux" = "/home/timvigers/OneDrive/Vigers/CF/Jordana Hoppe/SSwAB-IT"
)
knitr::opts_knit$set(root.dir = home_dir)
```

@@

  db_clinic_obtained_values %>% select(record_id, cos_collection_date, ppfev1),

@@

# Fix chronic oral antibiotics
db_medications$chronic_oral_antibiotic[
  db_medications$other_chronic_oral_antibiotic == "Azithromycin"
] <- "No"

@@

db_clinic_obtained_values <- left_join(db_clinic_obtained_values,

@@

```{r}
db_demographics %>%
  select(group, age_at_cos, age_group, ppfev1, sex, Race, Ethnicity) %>%
  tbl_summary(
    by = group, missing_text = "Missing",
    statistic = all_continuous() ~ c("{median} ({min}, {max})"),
    label = list(
      "sex" = "Sex", "age_at_cos" = "Age at COS", "age_group" = "Age Group"
    )
  ) %>%
  add_overall() %>%
  add_p(
    test = list(c("Race", "Ethnicity") ~ "fisher.test"),
    test.args = age_group ~ list(workspace = 2e9)
  ) %>%
  separate_p_footnotes() %>%
  modify_footnote(everything() ~ NA) %>%
  as_gt()
```

## By age group

```{r}
db_demographics %>%
  mutate(age_group = cut(age_at_cos, c(-Inf, 18, Inf), right = F)) %>%
  select(group, age_group, ppfev1, sex, Race, Ethnicity) %>%
  tbl_summary(
    by = age_group, missing_text = "Missing",
    statistic = all_continuous() ~ c("{median} ({min}, {max})"),
    label = list(
      "sex" = "Sex", "age_at_cos" = "Age at COS"
    )
  ) %>%
  add_overall() %>%
  add_p(
    test = list(c("Race", "Ethnicity") ~ "fisher.test")
  ) %>%
  separate_p_footnotes() %>%
  as_gt()

db_demographics %>%
  mutate(age_group = cut(age_at_cos, c(-Inf, 18, Inf), right = F)) %>%
  select(age_group, ppfev1) %>%
  drop_na() %>%
  ggplot(., aes(x = ppfev1)) +
  geom_histogram() +
  facet_wrap(~age_group)
```

@@

```{r}
# All pathogens
d_overall <- df[, c("overall_cos", "overall_sos")]
colnames(d_overall) <- c("COS", "SOS")
d_overall <- d_overall[complete.cases(d_overall), ]
a1 <- plot(euler(d_overall == "Positive"), quantities = list(cex = 3))
a2 <-
  ggplot(as.data.frame(table(d_overall$COS, d_overall$SOS)), aes(Var1, Var2)) +
  geom_tile(color = "white", fill = "#1E88E5") +
  geom_text(
    aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
    color = "white"
  ) +
  xlab("COS") +
  ylab("SOS") +
  theme_bw() +
  theme(text = element_text(size = 15))
# MRSA
d_MSSA <- df[, c("staphylococcus_aureus_cos", "staphylococcus_aureus_sos")]
colnames(d_MSSA) <- c("COS", "SOS")
d_MSSA <- d_MSSA[complete.cases(d_MSSA), ]
b1 <- plot(euler(d_MSSA == "Positive"), quantities = list(cex = 3))
b2 <-
  ggplot(as.data.frame(table(d_MSSA$COS, d_MSSA$SOS)), aes(Var1, Var2)) +
  geom_tile(color = "white", fill = "#1E88E5") +
  geom_text(aes(label = paste0(Freq, "\n(", round(Freq / sum(Freq), 3) * 100, "%)")),
    color = "white"
  ) +
  xlab("COS") +
  ylab("SOS") +
  theme_bw() +
  theme(text = element_text(size = 15))
p1 <- ggarrange(a1, a2, b1, b2, nrow = 2, ncol = 2) +
  bgcolor("White")
ggsave("./Dissemination/Figures/fig.png",
  plot = p1, width = 9, height = 6, units = "in"
)
a1 <- plot(euler(d_overall == "Positive"),
  quantities = list(cex = 3),
  main = "Any CF Pathogen"
)
b1 <- plot(euler(d_MSSA == "Positive"),
  quantities = list(cex = 3),
  main = "MSSA"
)
p2 <- ggarrange(a1, b1, nrow = 1, ncol = 2) +
  bgcolor("White")
ggsave("./Dissemination/Figures/euler_bw.png",
  plot = p2, width = 9, height = 6, units = "in"
)
a1 <- plot(venn(d_overall == "Positive"),
  quantities = list(cex = 3),
  main = "Any CF Pathogen"
)
b1 <- plot(venn(d_MSSA == "Positive"),
  quantities = list(cex = 3),
  main = "MSSA"
)
p3 <- ggarrange(a1, b1, nrow = 1, ncol = 2) +
  bgcolor("White")
ggsave("./Dissemination/Figures/venn_bw.png",
  plot = p3, width = 9, height = 6, units = "in"
)
```
